---
title: "Descriptive statistics"
author: "Erik Ø. Sørensen"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(targets)
library(tidyverse)
library(modelsummary)
library(gt)
```

```{r Loading data, include=FALSE}
tar_load(decisions_complete)
tar_load(players)
tar_load(players_complete)
tar_load(answers_complete)
tar_load(answersd)
```



## Sample and attrition


### Attrition
We want to know the extent of attrition and how it
differs by treatment. This we can tell from the "status" variable
on the players (post treatment assignment):
```{r Attrition, echo=FALSE}
players_assigned <- players |> filter(status>10)
n_status <- table(players_assigned$status, players_assigned$treatment)
chisq.test(n_status)
att_table <- prop.table(n_status, 2) |>
  as.data.frame() |> 
  pivot_wider(id_cols="Var1", names_from="Var2", values_from = "Freq") |>
  filter(! (Var1 %in% c(50,110))) 
att_table <- att_table |>
  mutate(across(where(is.numeric), ~ round(., 4))) |>
  bind_rows(
    att_table |>
      summarise(across(where(is.numeric), sum)) |>
      mutate(Var1 = "Total attrition:") |>
      relocate(Var1)
  ) |>
  mutate(Var1 = recode(Var1,
                       `20` = "Reading instructions",
                       `30` = "Making decisions",
                       `40` = "Answering background questions",
                       `100` = "Revoking consent")) |>
  rename(` ` = Var1,
        "Now"=now,
        "Short"=short,
        "Long"=long,
        "Never"=never)
attrition <- att_table |> gt() |> fmt_number(decimals=3) |> 
   tab_style(
    style = cell_borders(
      sides = "top",
      color = "black",
      weight = px(2)
    ),
    locations = cells_body(
      rows = ` ` == "Total attrition:"
    )
  )
attrition
gtsave(attrition, here::here("tables","attrition.tex"))
```


### Representativeness of sample

The data for the population in 2019 have been downloaded from 
[Statistics Norway, Table 10211](https://www.ssb.no/statbank/table/10211/).

```{r Representativeness of sample, echo=FALSE}
tar_load(representativeness_of_sample)
representativeness_of_sample$panelA 
gtsave(representativeness_of_sample$panelA , here::here("tables","representativeness_of_sampleA.tex"))
representativeness_of_sample$panelB
gtsave(representativeness_of_sample$panelA , here::here("tables","representativeness_of_sampleB.tex"))
```

### Descriptives on sample

```{r Descriptives on sample, echo=FALSE}
tar_load(descriptives_on_sample)
descriptives <- descriptives_on_sample |> gt::gt() |> fmt_number(decimals=3)
descriptives
gtsave(descriptives, here::here("tables", "descriptives_on_sample.tex"))
```




### Other numbers for text

What is the correlation between own risk preferences and beliefs about others?
```{r}
cor(answers_complete$risk_own, answers_complete$risk_others)
```


What about the timing of decisions? First, look at time to first decision:
```{r Time to first decision, echo=FALSE}
first_decisions <-decisions_complete %>% group_by(psid) %>% 
  summarize(first = min(updated_at)) %>%
  left_join(players_complete) %>%
  mutate( diff_time = first - created_at)
first_decisions %>%
  summarize(p05=quantile(diff_time, probs=0.05), 
            p25=quantile(diff_time, probs=0.25),
            p50=median(diff_time),
            p75=quantile(diff_time, probs=0.75),
            p95=quantile(diff_time, probs=0.95),
            avg = mean(diff_time),
            below60 = mean( (diff_time<60)))

```

Time to subsequent decisions:
```{r Time to subsequent decisions, echo=FALSE}
first_decisions <-decisions_complete %>% group_by(psid) %>% 
  summarize(first = min(updated_at)) %>%
  left_join(players_complete) %>%
  mutate( diff_time = first - created_at)
first_decisions %>%
  summarize(p05=quantile(diff_time, probs=0.05), 
            p25=quantile(diff_time, probs=0.25),
            p50=median(diff_time),
            p75=quantile(diff_time, probs=0.75),
            p95=quantile(diff_time, probs=0.95),
            avg = mean(diff_time),
            below60 = mean( (diff_time<60)))
```



## Decisions

```{r Distribution of choices, echo=FALSE}
tar_load(big_histogram_gg)
big_histogram_gg
ggsave(here::here("graphs","big_histogram.pdf"), width=24, height =16, units="cm")
```



```{r Graph of average outcomes, echo=FALSE}
dist_prop_lottery_choice <- decisions_complete |>
  group_by(psid) |>
  summarize( avgchoice = mean(choice_risk)) |>
  left_join(players_complete, by="psid") |>
  left_join(answers_complete, by="psid") |>
  ggplot() + 
  geom_histogram(aes(avgchoice, (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]),   
                 bins=29) + 
  facet_wrap(.~treatment) + theme_minimal() + ylab("Proportion") +
  xlab("Proportion of situations in which the lottery was chosen")
dist_prop_lottery_choice
ggsave(here::here("graphs","avgchoice_by_treatment.pdf"), width=16, height = 10, units="cm")


```
### Equality of distribution?

First, what are the means? 
```{r Means by treatment, echo=FALSE}
avgchoice <- decisions_complete |> 
  group_by(psid) |>
  summarize( avgchoice = mean(choice_risk)) |>
  left_join(players_complete, by="psid") |>
  left_join(answers_complete, by="psid")

avgchoice |> group_by(treatment) |>
  summarize(mean_avgchoice = mean(avgchoice),
            sd_avgchoice = sd(avgchoice),
            se_avgchoice = sd(avgchoice)/sqrt(n()),
            median_avgchoice = median(avgchoice)) |> gt::gt() |> fmt_number(decimals=3)
avgchoice |> 
  summarize(mean_avgchoice = mean(avgchoice),
            sd_avgchoice = sd(avgchoice),
            se_avgchoice = sd(avgchoice)/sqrt(n()),
            median_avgchoice = median(avgchoice)) |> gt::gt() |> fmt_number(decimals=3)
```

I use a 3-way extension of the Epps-Singleton test, implemented in R using the `manova` command (which
do the Hotelling-Lawley test, which is the K-way extension of the Hotelling test (equivalent to Epps-Singleton
on suitably defined variables).)

```{r Epps-Singleton-Hotelling test, echo=FALSE}
s <- IQR(avgchoice$avgchoice)
dist_test_df <- avgchoice |> mutate( g1 = cos(0.4*avgchoice/s),
                                   g2 = sin(0.4*avgchoice/s),
                                   g3 = cos(0.8*avgchoice/s),
                                   g4 = sin(0.8*avgchoice/s)) 
maov <- dist_test_df |>
  manova( cbind(g1,g2,g3,g4) ~ treatment, data=_) 
summary(maov, test = "Hotelling-Lawley")
```

### Inconsistencies
Participants can be said to act inconsistently if their choices over 
a lottery is not monotone in the value of the safe alternative. 


```{r Inconsistencies, echo=FALSE}
decisions_complete |> 
  group_by(psid,dienumber) |> 
  arrange(psid,dienumber,safe_amount) |>
  mutate(risk_seeking_switch = (choice_risk - lag(choice_risk))==1,
         safe_seeking_switch = (lag(choice_risk) - choice_risk)==1) |>
  summarize(left = max(risk_seeking_switch, na.rm=TRUE),
            right = max(safe_seeking_switch, na.rm=TRUE),
            all_safe = (max(choice_risk)==0),
            all_risk = (min(choice_risk)==1)) |>
  left_join(players_complete) |>
  group_by(treatment) |>
  summarize(share_allrisk = mean(all_risk),
            share_allsafe = mean(all_safe),
            share_inconsistent = mean(left),
            share_rat_switch = mean(right)) |>
  gt::gt() |> fmt_number(decimals = 3)
```
